---
title: "ASSIGNMENT 5"
author: "Simek, Kathryn"
date: '2020-09-20'
output:
  word_document: default
  html_notebook: default
  pdf_document: default
---
### Question 1: What are the elements in your data (including the categories and data types)?
#### Answer: 
* Id - quantitative - character
* Id2 - quantitative - integer
* Geography - categorical - character
* PopGroupId - quantitative - integer
* POPGROUP.display-label - categorical - character
* RacesReported - quantitative - integer
* HSDegree - quantitative - numeric
* BachDegree - quantitative - numeric

### Question 2. Please provide the output from the following functions: str(); nrow(); ncol()

```{r}
## Set the working directory to the root of your DSC 520 directory
setwd("C:/Users/katie/OneDrive/Documents/GitHub/dsc520")

## Load the ggplot2 package
library(ggplot2)
theme_set(theme_minimal())

acs_df <- read.csv("data/acs-14-1yr-s0201.csv")

```

### str():
```{r}
str(acs_df)
```

### nrow():
```{r}
nrow(acs_df)
```

### ncol():
```{r}
ncol(acs_df)
```

### Question 3. Create a Histogram of the HSDegree variable.
#### a. Set a bin size for the Histogram.
#### b. Include a Title and appropriate X/Y axis labels on your Histogram Plot.

```{r}
ggplot(acs_df, aes(HSDegree))+geom_histogram(bins = 68) +
      ggtitle("Likelyhood of High School Degree Per County") +
      xlab("Percent of Population with HS Degree") +
      ylab("Number of Counties")  
```

### Question 4. Answer the following questions based on the Histogram produced:
### a. Based on what you see in this histogram, is the data distribution unimodal?
#### Answer: Yes, the data distribution is unimodal.

### b. Is it approximately symmetrical?
#### Answer: No, the data distribution is not symmetrical.

### c. Is it approximately bell-shaped?
#### Answer: Yes, the data is approximately bell-shaped.

### d. Is it approximately normal?
#### Answer: No, it is not approximately normal.

### e. If not normal, is the distribution skewed? If so, in which direction?
#### Answer: The distribution is skewed negatively.

### f. Include a normal curve to the Histogram that you plotted.
```{r}
ggplot(acs_df, aes(HSDegree)) + geom_histogram(bins = 68) +
  ggtitle("Likelyhood of High School Degree Per County") +
  xlab("Percent of Population with HS Degree") +
  ylab("Number of Counties") + 
  stat_function(fun = dnorm, args = list(mean = mean(acs_df$HSDegree, 
                                                     na.rm = TRUE), 
                                         sd = sd(acs_df$HSDegree, 
                                                 na.rm = TRUE)))
```

### g. Explain whether a normal distribution can accurately be used as a model for this data.
#### Answer: No, the normal curve is significantly different from the probability plot. The plot is skewed quite a bit.


### Question 5. Create a Probability Plot of the HSDegree variable.
```{r}
ggplot(acs_df, aes(HSDegree))+geom_density() +
      ggtitle("Likelyhood of High School Degree Per County") +
      xlab("Percent of Population with HS Degree") +
      ylab("Number of Counties")
```

### Question 6. Answer the following questions based on the Probability Plot:
### a. Based on what you see in this probability plot, is the distribution approximately normal? Explain how you know.
#### Answer: The plot does not have a symetrical shape, so the distribution is not normal.

### b. If not normal, is the distribution skewed? If so, in which direction? Explain how you know.
#### Answer: The distribution is skewed negatively, the tail on the left side of the peak is longer and thinner than the right side of the curve.


### Question 7. Now that you have looked at this data visually for normality, you will now quantify normality with numbers using the stat.desc() function. Include a screen capture of the results produced.
```{r}
library(psych)
describe(acs_df)
```
```{r}
library(moments)
skewness(acs_df$HSDegree)
```
```{r}
library(moments)
kurtosis(acs_df$HSDegree)
```
```{r}
hsdegree_z <-(acs_df$HSDegree - mean(acs_df$HSDegree)) / sd(acs_df$HSDegree)
hsdegree_z
```

### Question 8. In several sentences provide an explanation of the result produced for skew, kurtosis, and z-scores. In addition, explain how a change in the sample size may change your explanation?
#### Answer: The skew of -1.69 is greater than a normal skew range of -1 to 1 so the skew is not normal, it is significantly negatively skewed. The kurtosis of 7.46 is greater than the -3 to 3 range so the kurtosis is not normal, it is significantly more positive, or leptokurtic. With the z-scores, we can see that most values are pretty significant, with a large range in values, which describes why we have heavy tails and many outliers in our distribution. 

  